Methods paper for erosion

### 2017-10-20 docs/methods/erosion_methods.tex
small changes to numerics

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align*}
  J[\eeta](\xx_0) = \frac{1}{2}\pderiv{\eeta}{\ss} \cdot \ss
  \left[ \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
\end{align*}
The deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{k=1}^{M}
    \frac{\rr \cdot \llambda_k}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{k=1}^M
    \frac{\xi_k}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right),
  \end{aligned}
\end{equation}
where $\rr = \xx - \cc_k$, $\cc_k$ is the point chosen inside body
$\gamma_k$, and $\rho = \|\rr_k\|$.  Now that the deformation tensor of
each component of the completed double-layer potential is computed, the
shear stress at $\xx_0 \in \bd\Omega$ is
\begin{align*}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Pressure}
To compute the pressure at $\xx \in \overline{\Omega}$, we follow the
same procedure used for the shear stress.  We start by computing the
pressure of the double-layer potential for $\xx \in \Omega$ and add the
appropriate jump if $\xx \in \Gamma$. For $\xx \in \Omega$, the pressure
is
\begin{align}
  p^{D}(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta(\yy) \,ds_\yy.
    \label{eqn:pressureDLP}
\end{align}
To compute the pressure at $\xx_0 \in \bd\Omega$, we consider the limit
as $\xx$ tends to $\xx_0$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
    \pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0).
\end{align*}
The pressure due to the Stokeslets is
\begin{align*}
  p^S(\xx_0) = \sum_{k=1}^{M}\frac{\rr \cdot \llambda_k}{2\pi\rho^2},
\end{align*}
and the pressure due to the rotlets is zero.  Therefore, the pressure is
\begin{align*}
  p(\xx) &= p^{\DD}(\xx) + p^{S}(\xx), \quad &&\xx \in \Omega,\\
  p(\xx) &= \pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx) + 
              p^{S}(\xx), && \xx \in \bd\Omega.
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Shear Stress}
We have decomposed the deformation tensor into the contributions due to
three different terms.  The contributions from the Stokeslets and the
rotlets can be directly evaluated since $\rr \neq 0$ for all $\xx \in
\gamma$.  However, the kernel of the deformation tensor due to the
double-layer potential~\eqref{eqn:shearStressDLP} 
\begin{align*}
  K_\ssigma(\xx,\yy) = \frac{1}{2\pi} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) +
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right),
\end{align*}
has a singularity that is asymptotic to $\bigO(\rho^{-2})$ when $\yy
\rightarrow \xx$.  The singularity strength can be reduced since the
velocity corresponding to a constant density function is
constant~\cite{poz1992}.  Therefore, the deformation tensor of the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx) = \int_{\bd\Omega}K_\ssigma(\xx,\yy)
      (\eeta(\yy) - \eeta(\xx)) \, ds_{\yy}, \quad \xx \in \bd\Omega.
\end{align*}
This expression for the deformation tensor has a singularity that is
asymptotic to $\bigO(\rho^{-1})$ when $\yy \rightarrow \xx$.  Finally,
to achieve spectral accuracy, we apply odd-even
integration~\cite{sid-isr1988}.  That is, the deformation tensor due to
the double-layer potential at a discretization point $\xx_i$ on body
$\gamma_k$, with $i$ being odd, is approximated as
\begin{align}
  \ssigma^{\DD}(\xx_i) = \sum_{\xx_j \notin \gamma_k}
    K_\ssigma(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_j + 
  \sum_{\substack{\xx_j \in \gamma_k \\ j \: \mathtt{even}}}
    K_\ssigma(\xx_i,\xx_j) (\eeta(\xx_j) - \eeta(\xx_i)) \Delta s_j,
  \label{eqn:stressOddEven}
\end{align}
and a similar approximation is used when $i$ is even.  The deformation
tensor due to the Stokeslets and rotlets is given in
equation~\eqref{eqn:shearStressSR}, and then the shear stress can be
computed by computing the deformation tensor in the normal direction,
and taking the inner product with the tangent direction.

\todo[inline]{discuss how we handle jump condition ... Fourier
differentiation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Pressure}
Similar to the shear stress, the singularity of the pressure
corresponding to the double-layer potential~\eqref{eqn:pressureDLP} is
too strong for the trapezoid rule to be directly applied.  Therefore, we
again subtract a constant density function and write the pressure as
\begin{align*}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot (\eeta(\yy) - \eeta(\xx_0)) \,ds_\yy.
\end{align*}
Then, identical to how the shear stress is
computed~\eqref{eqn:stressOddEven}, odd-even integration is used to
evaluate this layer potential for pressure.

\todo[inline]{discuss how we handle jump condition ... Fourier
differentiation}

>>>>

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\pderiv{\eeta}{\ss} \cdot \ss
  \left[ \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
The deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{k=1}^{M}
    \frac{\rr \cdot \llambda_k}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{k=1}^M
    \frac{\xi_k}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right),
  \end{aligned}
\end{equation}
where $\rr = \xx - \cc_k$, $\cc_k$ is the point chosen inside body
$\gamma_k$, and $\rho = \|\rr_k\|$.  Now that the deformation tensor of
each component of the completed double-layer potential is computed, the
shear stress at $\xx_0 \in \bd\Omega$ is
\begin{align*}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Pressure}
To compute the pressure at $\xx \in \overline{\Omega}$, we follow the
same procedure used for the shear stress.  We start by computing the
pressure of the double-layer potential for $\xx \in \Omega$ and add the
appropriate jump if $\xx \in \Gamma$. For $\xx \in \Omega$, the pressure
is
\begin{align}
  p^{D}(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta(\yy) \,ds_\yy.
    \label{eqn:pressureDLP}
\end{align}
To compute the pressure at $\xx_0 \in \bd\Omega$, we consider the limit
as $\xx$ tends to $\xx_0$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
    \pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0).
  \label{eqn:pressureJump}
\end{align}
The pressure due to the Stokeslets is
\begin{align*}
  p^S(\xx_0) = \sum_{k=1}^{M}\frac{\rr \cdot \llambda_k}{2\pi\rho^2},
\end{align*}
and the pressure due to the rotlets is zero.  Therefore, the pressure is
\begin{align*}
  p(\xx) &= p^{\DD}(\xx) + p^{S}(\xx), \quad &&\xx \in \Omega,\\
  p(\xx) &= \pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx) + 
              p^{S}(\xx), && \xx \in \bd\Omega.
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Shear Stress}
We have decomposed the deformation tensor into the contributions due to
three different terms.  The contributions from the Stokeslets and the
rotlets can be directly evaluated since $\rr \neq 0$ for all $\xx \in
\gamma$.  However, the kernel of the deformation tensor due to the
double-layer potential~\eqref{eqn:shearStressDLP} 
\begin{align*}
  K_\ssigma(\xx,\yy) = \frac{1}{2\pi} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) +
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right),
\end{align*}
has a singularity that is asymptotic to $\bigO(\rho^{-2})$ when $\yy
\rightarrow \xx$.  The singularity strength can be reduced since the
velocity corresponding to a constant density function is
constant~\cite{poz1992}.  Therefore, the deformation tensor of the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx) = \int_{\bd\Omega}K_\ssigma(\xx,\yy)
      (\eeta(\yy) - \eeta(\xx)) \, ds_{\yy}, \quad \xx \in \bd\Omega.
\end{align*}
This expression for the deformation tensor has a singularity that is
asymptotic to $\bigO(\rho^{-1})$ when $\yy \rightarrow \xx$.  Finally,
to achieve spectral accuracy, we apply odd-even
integration~\cite{sid-isr1988}.  That is, the deformation tensor due to
the double-layer potential at a discretization point $\xx_i$ on body
$\gamma_k$, with $i$ being odd, is approximated as
\begin{align}
  \ssigma^{\DD}(\xx_i) = \sum_{\xx_j \notin \gamma_k}
    K_\ssigma(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_j + 
  \sum_{\substack{\xx_j \in \gamma_k \\ j \: \mathtt{even}}}
    K_\ssigma(\xx_i,\xx_j) (\eeta(\xx_j) - \eeta(\xx_i)) \Delta s_j,
  \label{eqn:stressOddEven}
\end{align}
and a similar approximation is used when $i$ is even.  The deformation
tensor due to the Stokeslets and rotlets is given in
equation~\eqref{eqn:shearStressSR}.  Finally, the jump in the
deformation tensor~\eqref{eqn:shearStressJump} is added, and the
required arclength derivative is computed in Fourier space.  Then, the
shear stress can be computed by computing the deformation tensor in the
normal direction, and taking the inner product with the tangent
direction.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Pressure}
Similar to the shear stress, the singularity of the pressure
corresponding to the double-layer potential~\eqref{eqn:pressureDLP} is
too strong for the trapezoid rule to be directly applied.  Therefore, we
again subtract a constant density function and write the pressure as
\begin{align*}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot (\eeta(\yy) - \eeta(\xx_0)) \,ds_\yy.
\end{align*}
Then, identical to how the shear stress is
computed~\eqref{eqn:stressOddEven}, odd-even integration is used to
evaluate this layer potential for pressure.  Finally, to add the jump
condition~\eqref{eqn:pressureJump}, Fourier differentiation is used to
compute the arclength derivative.
